\bf Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-04-21 paper.qmd
Update the algorithm section

When $k=2$ which is assumed for hexagon binning, $y_{1,\text{max}}=1$ and $y_{2,\text{max}} = \frac{r_2}{r_1}\times \frac{2}{\sqrt{3}}$. (The $\frac{2}{\sqrt{3}}$ accounts for the different height ($h_b$) and width ($w_b$) of a regular hexagon.)

@@

To define a hexagonal grid across the \kD{} space, using hexagons with fixed height ($h_b$), width ($w_b$), need to compute the how many hexagons should represented by each axes in \kD{} space.

When $k=2$, the number of bins along the $x$ and $y$ axes, $b_1$ and $b_2$ is computed by accounting ${r_1, r_2}$, offset along the axes ($q_1$, $q_2$), and height ($h_b$), width ($w_b$) of a regular hexagon.

($b_1 = \frac{r_1 + q_1}{w_b}$, $b_2 = \frac{r_2 + q_2}{0.75 \times h_b}$): $0.75 \times h_b$ acommodate to have the tesselation without spacing.

@@

Define a hexagonal grid across the \kD{} space, using hexagons with fixed height ($h_b$), width ($w_b$).

When $k=2$,

The starting coordinates $(s_1, s_2)$ mark the lower left of the grid. This is the bottom left bin centroid. By starting from there, points are generated to fill the grid accounting $b_1$ and $b_2$.

<!--We deliberately separate out the creation of the hexagonal grid from the mapping of points on the grid.-->
Each of the \kD{} points will belong to a hexagonal bin. That is, for each $y \in \mathbfit{Y}$, we can (uniquely) identify the hexagon that the point belongs to. This identification is done finding the nearest bin centroid for the \kD{} points by considering the \kD{} Euclidean distance.

#### Obtaining bin centroids

Let $h : \mathbb{R}^{n\times k} \rightarrow \mathbb{R}^{m\times k}$ be the function that maps a point in \kD{} space to it's hexagonal centroid ($C^{(k)}$).

@@

computes a set of simplices such that no data points of X are contained in any circumspheres of the simplices.

When $k = 2$

Delaunay triangulation of $C^{(k)}$ is a triangulation of the convex hull of $C^{(k)}$ such that the circumcircle of every triangle in the triangulation contains no other points from $C^{(k)}$.

@@

When $k=2$. which bin centroids should connected are identified as a result of triangulation. Therefore, corresponding \pD{} points should connected in the same order.

>>>>

When $k=2$ which is assumed for hexagon binning, $y_{1,\text{max}}=1$ and $y_{2,\text{max}} = \frac{r_2}{r_1}\times \frac{2}{\sqrt{3}}$. (The $\frac{2}{\sqrt{3}}$ accounts for the different height ($h_b$) and width ($w_b$) of a regular hexagon.) (@fig-NLDR-scurve (a))

@@

To define a hexagonal grid across the \kD{} space, with hexagons of fixed height ($h_b$), width ($w_b$),  it is necessary to determine how many hexagons should be represented along each axis in the \kD{} space.

When $k=2$, the number of bins along the $x$ and $y$ axes, $b_1$ and $b_2$, is computed by considering the scaling factors ${r_1, r_2}$, along with the offset along the axes ($q_1$, $q_2$), and the height ($h_b$) and width ($w_b$) of a regular hexagon (@fig-NLDR-scurve (b)). ($b_1 = \frac{r_1 + q_1}{w_b}$, and $b_2 = \frac{r_2 + q_2}{0.75 \times h_b}$) where $0.75 \times h_b$ accommodate to have hexagons fill the entire 2D space without leaving any gaps between them.)

@@

Define a hexagonal grid across the \kD{} space, using hexagons with fixed height ($h_b$), width ($w_b$). Each of the \kD{} points will belong to a hexagonal bin. That is, for each $y \in \mathbfit{Y}$, we can (uniquely) identify the hexagon that the point belongs to. This identification is done finding the nearest bin centroid for the \kD{} points by considering the \kD{} Euclidean distance.

When $k=2$, the starting coordinates $(s_1, s_2)$ mark the lower left of the grid. This is the bottom left bin centroid. By starting from there, points are generated to fill the grid accounting $b_1$, $b_2$, $w_b$, and $h_b$.

<!--We deliberately separate out the creation of the hexagonal grid from the mapping of points on the grid.-->
#### Obtaining bin representations

Let $h : \mathbb{R}^{n\times k} \rightarrow \mathbb{R}^{m\times k}$ be the function that maps a point in \kD{} space to it's bin representation ($C^{(k)}$). This is the model points in \kD{} space.

When $k = 2$, one of the representations of a bin is the hexagonal centroid (@fig-NLDR-scurve (c)).

@@

Delaunay triangulation on bin representations generates edges by specifying which bin representations should be connected to generate the model in \kD{} space.

When $k = 2$ Delaunay triangulation on $C^{(2)}$ generates the model in 2D space, which is a triangular mesh (@fig-NLDR-scurve (d)). It generates convex hulls of $C^{(2)}$ such that the circumcircle of every triangle in the triangulation contains no other points from $C^{(2)}$.

@@

Furthermore, edges that exist between \kD{} representations should also generate edges in \pD{} by connecting \pD{} mapping of the corresponding \kD{} representations.

### 2024-04-21 paper.tex
scaling data...right
Update the algorithm section

\begin{tabular}{>{\raggedright\arraybackslash}p{3cm}>{\raggedright\arraybackslash}p{12cm}}
\toprule
\textbf{Notation} & \textbf{Description}\\
\midrule
$n, p, k, m$ & number of observations, variables, embedding dimension, number of non-empty bins, respectively\\
$\mathbfit{X}, \mathbfit{x}$ & $p$-dimensional data (population, sample)\\
$\mathbfit{y}$ & $k$-dimensional layout\\
$P$ & orthonormal basis, generating a $d\text{-}dimensional$ linear projection of $p$-dimensional data\\
$T$ & true  model\\
\addlinespace
$g$ & functional mapping from \pD{} to \kD{}, especially as prescribed by NLDR\\
$\mathbfit{\theta}$ & (Hyper-) parameters for NLDR method\\
$y_\text{max}$ & maximum value of the scaled second embedding component\\
$r$ & ranges of the embedding components\\
$b$ & bins\\
\addlinespace
$a_r$ & aspect ratio of the 2D layout before scaling\\
$s$ & hexagonal size (defined as the radius of the outer circle surrounding the hexagon)\\
$(x_s, \ y_s)$ & starting coordinates of the hexagonal grid\\
$q_x$, $q_y$ & buffer amount along the x and y axes\\
$h_s$, $v_s$ & horizontal and vertical distance between adjacent hexagon bin centroids\\
\addlinespace
$d_x$, $d_y$ & horizontal spacing and adjusted vertical spacing for hexagonal coordinates\\
$b_x$, $b_y$ & number of hexagon bins along the x and y axes\\
$b$, $b'$ & total and non-empty hexagon bins in the grid\\
$n_k$ & number of observations within the $k^{th}$ hexagon\\
$(h^k_{x_i}, h^k_{y_i})$ & hexagonal grid coordinates of the $k^{th}$ hexagon\\
\addlinespace
$C^{(j)}$ & $j$-dimensional bin centers\\
\bottomrule
\end{tabular}
\endgroup{}

@@

When \(k=2\) which is assumed for hexagon binning,
\((y_{1,\text{max}}=1\) and
\(y_{2,\text{max}} = \frac{r_2}{r_1}\times \frac{2}{\sqrt{3}}\). (The
\(\frac{2}{\sqrt{3}}\) accounts for the different height and width of a
regular hexagon.)

@@

Define a hexagonal grid across the \kD{} space, using hexagons with
fixed height (\(h_b\)), width (\(w_b\)). The starting coordinates
\((x_s, y_s)\) mark the lower left of the grid. Each of the \kD{} points
will belong to a hexagonal bin. That is, for each
\(y \in \mathbfit{Y}\), we can (uniquely) identify the hexagon that the
point belongs to.

\subsubsection{Obtaining bin centroids}\label{obtaining-bin-centroids}

Let \(h : \mathbb{R}^{n\times k} \rightarrow \mathbb{R}^{m\times k}\) be
the function that maps a point in \kD{} space to it's hexagonal centroid
(\(C^{(k)}\)).

\subsubsection{Triangulating centroids}\label{triangulating-centroids}

@@

>>>>

\begin{tabular}{>{\raggedright\arraybackslash}p{3cm}>{\raggedright\arraybackslash}p{12cm}}
\toprule
\textbf{Notation} & \textbf{Description}\\
\midrule
$n, p, k, m$ & number of observations, variables, embedding dimension, number of non-empty bins, respectively\\
$\mathbfit{X}, \mathbfit{x}$ & $p$-dimensional data (population, sample)\\
$\mathbfit{y}$ & $k$-dimensional layout\\
$P$ & orthonormal basis, generating a $d\text{-}dimensional$ linear projection of $p$-dimensional data\\
$T$ & true  model\\
\addlinespace
$g$ & functional mapping from \pD{} to \kD{}, especially as prescribed by NLDR\\
$\mathbfit{\theta}$ & (Hyper-) parameters for NLDR method\\
$y_\text{max}$ & maximum value of the scaled second embedding component\\
$r$ & ranges of the embedding components\\
$b$ & bins\\
\addlinespace
$a_r$ & aspect ratio of the 2D layout before scaling\\
$s$ & hexagonal size (defined as the radius of the outer circle surrounding the hexagon)\\
$(s_1, \ s_2)$ & starting coordinates of the hexagonal grid\\
$q_1$, $q_2$ & buffer amount along the x and y axes\\
$h_s$, $v_s$ & horizontal and vertical distance between adjacent hexagon bin centroids\\
\addlinespace
$d_x$, $d_y$ & horizontal spacing and adjusted vertical spacing for hexagonal coordinates\\
$b_1$, $b_2$ & number of hexagon bins along the x and y axes\\
$b$, $b'$ & total and non-empty hexagon bins in the grid\\
$n_k$ & number of observations within the $k^{th}$ hexagon\\
$(h^k_{x_i}, h^k_{y_i})$ & hexagonal grid coordinates of the $k^{th}$ hexagon\\
\addlinespace
$C^{(j)}$ & $j$-dimensional bin centers\\
\bottomrule
\end{tabular}
\endgroup{}

@@

When \(k=2\) which is assumed for hexagon binning,
\(y_{1,\text{max}}=1\) and
\(y_{2,\text{max}} = \frac{r_2}{r_1}\times \frac{2}{\sqrt{3}}\). (The
\(\frac{2}{\sqrt{3}}\) accounts for the different height (\(h_b\)) and
width (\(w_b\)) of a regular hexagon.) (Figure~\ref{fig-NLDR-scurve}
(a))

\subsubsection{Computing hexagonal grid
configurations}\label{computing-hexagonal-grid-configurations}

To define a hexagonal grid across the \kD{} space, with hexagons of
fixed height (\(h_b\)), width (\(w_b\)), it is necessary to determine
how many hexagons should be represented along each axis in the \kD{}
space.

When \(k=2\), the number of bins along the \(x\) and \(y\) axes, \(b_1\)
and \(b_2\), is computed by considering the scaling factors
\({r_1, r_2}\), along with the offset along the axes (\(q_1\), \(q_2\)),
and the height (\(h_b\)) and width (\(w_b\)) of a regular hexagon
(Figure~\ref{fig-NLDR-scurve} (b)). (\(b_1 = \frac{r_1 + q_1}{w_b}\),
and \(b_2 = \frac{r_2 + q_2}{0.75 \times h_b}\)) where
\(0.75 \times h_b\) accommodate to have hexagons fill the entire 2D
space without leaving any gaps between them.)

@@

Define a hexagonal grid across the \kD{} space, using hexagons with
fixed height (\(h_b\)), width (\(w_b\)). Each of the \kD{} points will
belong to a hexagonal bin. That is, for each \(y \in \mathbfit{Y}\), we
can (uniquely) identify the hexagon that the point belongs to. This
identification is done finding the nearest bin centroid for the \kD{}
points by considering the \kD{} Euclidean distance.

When \(k=2\), the starting coordinates \((s_1, s_2)\) mark the lower
left of the grid. This is the bottom left bin centroid. By starting from
there, points are generated to fill the grid accounting \(b_1\),
\(b_2\), \(w_b\), and \(h_b\).

\subsubsection{Obtaining bin
representations}\label{obtaining-bin-representations}

Let \(h : \mathbb{R}^{n\times k} \rightarrow \mathbb{R}^{m\times k}\) be
the function that maps a point in \kD{} space to it's bin representation
(\(C^{(k)}\)). This is the model points in \kD{} space.

When \(k = 2\), one of the representations of a bin is the hexagonal
centroid (Figure~\ref{fig-NLDR-scurve} (c)).

\subsubsection{Generating edges}\label{generating-edges}

Delaunay triangulation on bin representations generates edges by
specifying which bin representations should be connected to generate the
model in \kD{} space.

When \(k = 2\) Delaunay triangulation on \(C^{(2)}\) generates the model
in 2D space, which is a triangular mesh (Figure~\ref{fig-NLDR-scurve}
(d)). It generates convex hulls of \(C^{(2)}\) such that the
circumcircle of every triangle in the triangulation contains no other
points from \(C^{(2)}\).

@@

Furthermore, edges that exist between \kD{} representations should also
generate edges in \pD{} by connecting \pD{} mapping of the corresponding
\kD{} representations.
